### 2018-05-25 intro.tex
Fix syntax errors according to the first revision by Prof. Xue

\section{Introduction}\label{sec:intro}
Dense linear algebra libraries are the most fundamental software in
scientific and engineering computing domains.
Basic Linear Algebra Subprograms (BLAS) (FIXME) defines a collection
of APIs which act as standard building blocks for dense matrix operations.
BLAS routines are divided into 3 levels,
level-1 for vector-vector operations,
level-2 for matrix-vetor operations,
and level-3 for matrix-matrix operations.
Processor vendors often provide BLAS implementations
highly optimized for their processors,
such as Intel MKL, AMD ACML and NVIDIA cuBLAS.
The HPC community have also contributed several high-quality
BLAS implementations, e.g. ATLAS~\cite{atlas},
GotoBLAS~\cite{gotoblas}, OpenBLAS~\cite{openblas},
and BLIS~\cite{blis}\cite{blisport}.

Among three levels, level-3 provides the most opportunities
for optimization because it has the highest computational complexity ($O(n^3)$).
Among all level-3 operations, GEneral Matrix Multiply (GEMM) is
of the most interest because all other level-3 operations can be defined
in terms of GEMM and some level-1 and level-2 operations.
Much effort have been spent on optimizing GEMM for different architectures,
both in general algorithm and in architecture specific ways.
(FIXME citations)

GEMM performs a matrix-multiply-accumulate operation.
Generally, an optimized GEMM implementation consists of two components,
(1) a highly optimized kernel routine to accomplish the computation, and
(2) an overall strategy to partition the workload into small tasks
and schedule these tasks to be executed effectively on the target processor.
The kernel routine is a serial program performing matrix-multiply-accumulate
on matrices from a single task.
The overall strategy partitions the workload by tiling the loops,
and determines a task schedule by choosing a specific traversal order
in the loop nest's iteration space.
In multi-threaded context, different tasks (loop tiles) are assigned to
different threads to parallelize the whole operation.
The main target of GEMM optimization is maximizing the floating-point
operation throughput, measured by FLoating-point Operations Per Second ($flops$).
On one hand, optimization of the serial kernel routine concentrates on
instruction scheduling to improve instruction throughput.
On the other hand, the overall strategy is responsible for optimizing
memory accesses to protect the kernel from data hunger.
In this paper, we focus on optimizing the memory access of GEMM.

Currently, two techniques, software prefetching and data packing,
are widely adopted in GEMM implementations to speed up memory accesses.
Software prefetch instructions are utilized
to load data into cache ahead of use, so memory accesses will not
cause long latencies and floating-point instructions can execute at peak throughput.
To better exploit the capability of on-chip caches,
matrix blocks the kernel operates on will be packed into continous memory buffer
whose footprint is within the cache capacity before the kernel routine is called.
Because GEMM is a computational intensive operation whose
arithmetic complexity is $O(n^3)$ and memory complexity is $O(n^2)$,
the overhead caused by data packing is negligible when
the input matrices are large~\cite{gotogemm}.

The software prefetching and data packing approach meets a challenge
when working on architectures with non-LRU shared caches.
In traditional high-performance processors, e.g. Intel Xeon series,
both L1 and L2 caches are private to processor cores
and LRU replacement policy is adopted.
Without considering the last level cache (LLC) if presented,
threads running on different cores prefetch their data into
different caches, and private data of different threads
cannot cause conflict cache misses.
This is not the case if caches are shared by several processor cores.
There are chances that a cache line prefetched by one thread
get evicted out before its lifetime ends because another thread prefetches
another cache line into the same cache set.
A simple solution would be shrinking the size of packed matrix
so data used by several threads can simultaneously live in the shared cache.
Due to the nature of set-associative caches,
this solution cannot completely eliminate inter-thread cache conflicts,
especially if the shared cache has a non-LRU replacement policy.
Evaluation results show that inter-thread cache conflicts on non-LRU shared caches
can heavily harm GEMM performance,
even with packed matrices shrinked to fit the cache size.

Following the trend in recent decades, modern architecture design
are introducing more and more cores on a single processor chip.
To reduce cost of on-chip cache memories and coherence networks,
shared cache may become common in future many-core processors.
And non-LRU replacement policies, for example, psudo-random,
may also be used to further reduce the cache design complexity.
As far as we know, some high-performance processors based on ARM technology,
have already adopted this design. (FIXME list the processors)
As a consequence, a better solution to the inter-thread cache conflicts problem
on non-LRU shared caches is meaningful.

In this paper, we present a Shared Cache Partition (SCP) method,
to reduce inter-thread cache conflicts on architectures with non-LRU share caches.
The key idea is partition the share cache into physically disjoint sets
and assign different sets to different threads. This can be achieved by
exploiting the memory mapping mechanisim of set-associative caches.

The main contributions of this paper are:
\begin{itemize}
\item We present a quantitative analysis of the negative effect of inter-thread cache
  conflicts on GEMM performance.
\item We propose SCP, a method for solving the inter-thread cache conflicts
  problem on architectures with non-LRU shared caches.
\item We implement SCP in the OpenBLAS library and evaluate it on Phytium 2000+,
  an emerging high-performance processor based on ARM's AArch64 architecture,
  which has 64 cores, private LRU L1 caches and psudo-random shared L2 caches.
  Evaluation results show that GEMM performance get improved consistently
  under various parallelism configurations, by an amount of $2.75\%$--$6.91\%$..
\end{itemize}

>>>>

\section{Introduction}\label{sec:intro}
Dense linear algebra libraries are the most fundamental software in
scientific and engineering computing domains.
Basic Linear Algebra Subprograms (BLAS) (FIXME) defines a collection
of APIs which act as standard building blocks for dense matrix operations.
BLAS routines are divided into 3 levels,
level-1 for vector-vector operations,
level-2 for matrix-vetor operations,
and level-3 for matrix-matrix operations.
Processor vendors often provide BLAS implementations
highly optimized for their processors,
such as Intel MKL, AMD ACML and NVIDIA cuBLAS.
The HPC community has also contributed several high-quality
BLAS implementations, e.g. ATLAS~\cite{atlas},
GotoBLAS~\cite{gotoblas}, OpenBLAS~\cite{openblas},
and BLIS~\cite{blis}\cite{blisport}.

Among the three levels of BLAS routines, level-3 provides the most opportunities
for optimization because it has the highest computational complexity ($O(n^3)$).
Among all the level-3 operations, GEneral Matrix Multiply (GEMM) is
of the most interest because all other level-3 operations can be defined
in terms of GEMM and some level-1 and level-2 operations.
Much effort has been spent on optimizing GEMM for different architectures.
%% both in general algorithm and in architecture specific ways.
(FIXME citations)

GEMM performs a matrix-multiply-accumulate operation.
Generally, an optimized GEMM implementation consists of two components,
(1) a highly optimized kernel routine to accomplish the computation, and
(2) an overall strategy to partition the workload into small tasks
and schedule these tasks to be executed effectively on the target processor.
The kernel routine is a serial program performing matrix-multiply-accumulate
on matrices from a single task.
The overall strategy partitions the workload by tiling the loops,
and determines a task schedule by choosing a specific traversal order
in the loop nest's iteration space.
In a multi-threaded context, different tasks (loop tiles) are assigned to
different threads to parallelize the whole operation.
The main target of GEMM optimization is maximizing the floating-point
operation throughput, measured by FLoating-point Operations Per Second ($flops$).
On one hand, optimization of the serial kernel routine concentrates on
instruction scheduling to improve instruction throughput.
On the other hand, the overall strategy is responsible for optimizing
memory accesses to protect the kernel from being idling waiting for data.
In this paper, we focus on optimizing the memory access of GEMM.

Currently, two techniques, software prefetching and data packing,
are widely adopted in GEMM implementations to speed up memory accesses.
Software prefetch instructions are utilized
to load data into cache before they are used, so memory accesses will not
cause long latencies and floating-point instructions can execute at peak throughput.
To better exploit the capability of on-chip caches,
matrix blocks the kernel operates on will be packed into continous memory buffer
whose footprint is within the cache capacity before the kernel routine is called.
Because GEMM is a computational intensive operation whose
arithmetic complexity is $O(n^3)$ and memory complexity is $O(n^2)$,
the overhead caused by data packing is negligible when
the input matrices are large~\cite{gotogemm}.

The software prefetching and data packing approach meet a challenge
when working on architectures with non-LRU shared caches.
In traditional high-performance processors, e.g. Intel Xeon series,
both L1 and L2 caches are private to processor cores
and LRU replacement policy is adopted.
Without considering the last level cache (LLC),
threads running on different cores prefetch their data into
different caches, and private data of different threads
cannot cause conflict cache misses.
This is not the case if caches are shared by several processor cores.
There are chances that a cache line prefetched by one thread
get evicted out before its lifetime ends because another thread prefetches
another cache line into the same cache set.
A simple solution would be shrinking the size of packed matrix
so data used by several threads can simultaneously reside in the shared cache.
Due to the nature of set-associative caches,
this solution cannot completely eliminate inter-thread cache conflicts,
especially if the shared cache has a non-LRU replacement policy.
Evaluation results show that inter-thread cache conflicts on non-LRU shared caches
can heavily harm GEMM performance,
even with packed matrices shrinked to fit int the cache.

Following the trend in recent decades, modern architecture design
is introducing more and more cores on a single processor chip.
To reduce cost of on-chip cache memories and coherence networks,
shared cache may become common in future many-core processors.
And non-LRU replacement policies, for example, psudo-random,
may also be used to further reduce the cache design complexity.
Indeed, some high-performance processors based on ARM technology,
have already adopted this design. (FIXME list the processors)
As a consequence, a better solution to the inter-thread cache conflicts problem
on non-LRU shared caches is meaningful.

In this paper, we present a Shared Cache Partitioning (SCP) method,
to reduce inter-thread cache conflicts on architectures with non-LRU share caches.
The key idea is partition the share cache into physically disjoint sets
and assign different sets to different threads. This can be achieved by
exploiting the memory mapping mechanisim of set-associative caches.

The main contributions of this paper are:
\begin{itemize}
\item We present a quantitative analysis of the negative effect of inter-thread cache
  conflicts on GEMM performance.
\item We propose SCP, a method for solving the inter-thread cache conflicts
  problem on architectures with non-LRU shared caches.
\item We implement SCP in the OpenBLAS library and evaluate it on Phytium 2000+,
  an emerging high-performance processor based on ARM's AArch64 architecture,
  which has 64 cores, private LRU L1 caches and psudo-random shared L2 caches.
  Evaluation results show that GEMM performance improves consistently
  under various parallelism configurations, by an amount of $2.75\%$--$6.91\%$..
\end{itemize}

### 2018-05-25 scp.tex
Fix syntax errors according to the first revision by Prof. Xue

This section demonstrates the SCP method to handle inter-thread cache conflicts.
The basic idea is that inter-thread cache conflicts can be eliminated
if packed matrices used by different threads are fetched to different cache sets.
In Section~\ref{subsec:example}, we demonstrate SCP by an example.
Then formal description is given in Section~\ref{subsec:formal}.

@@

Suppose we are implementing a double precision GEMM (DGEMM) on
the 4-core processor, i.e. $es=8B$.
We now determine the tiling factors used in DGEMM following the procedure described
by constraints (\ref{eq:constraints.reg})--(\ref{eq:constraints.l3}).
Various configurations for ($M_r$, $N_r$) can satisfy (\ref{eq:constraints.reg}),
e.g. $(4,8)$, $(8,4)$, $(6,8)$, and $(8,6)$. 
It takes careful consideration to determine which one is the best,
and a tuning process may be needed.
After tuning we got the best values $M_r = 4$ and $N_r = 8$.
By (\ref{eq:constraints.l1}),
$K_c \le \lfloor c_{L1}/es/nt_{L1}/(N_r + 2 M_r) \rfloor = \lfloor 32KB/8B/1/16 \rfloor = 256$.
Here we choose $K_c=256$.
By (\ref{eq:constraints.l2}),
$M_c \le \lfloor (c_{L2}/es/nt_{L2} - 2 N_r K_c )/ K_c \rfloor =
\lfloor (2MB/8B/4 - 2*8*256)/256 \rfloor = 240$.
To leave space for other data structures, as well as the program code,
it is reasonable to shrink $M_c$ to a slightly smaller value,
e.g. $192$, $208$ and $224$ are all proper candidates for $M_c$.
Similar to $M_r$ and $N_r$, a tunning process may be needed to
find the best value for $M_c$.
Here we choose $M_c = 192$.
As there is no L3 cache, (\ref{eq:constraints.l3}) need not to be
concerned and $N_c$ can be given a large value.
Here we choose $N_c = 1024NT = 4096$.
With tiling factors determined, the sizes of packed matrices
can be calculated. Table.~\ref{tab:msizes} list the sizes of packed matrices.

@@

Figure.~\ref{fig:partition.conventional} and Figure.~\ref{fig:partition.segmented}
essentially represent two different partition style of the shared cache,
refered as way-partition and set-partition, respectively.
To guarantee the data layout in set-partition,
$A_2$ cannot be stored in a single continous buffer any more.
Instead, $A_2$ is distributed to 12 continous memory segments, each of size $32KB$.
Distance from one segment to the next is equal to $wc_2=128KB$.
%% There is no memory fragment because the segment size is a multiple of
%% size of $A_4$ (8KB).
%% FIXME put this to the end of this section
While the set-partition of shared caches avoids inter-thread cache conflicts,
it introduces some complexity to GEMM implementation
because the GEBP kernel and the matrix packing routines
now must work with segmented memory buffers instead of continous ones.

How about $B_2$, the other packed matrix use in GEBP?
Unlike $A_2$, which is thread private data,
$B_2$ is shared among all threads and
the packing of $B_2$ is done collabratively by all threads.
We have two choices, privatize $B_2$ and apply the set-partition method,
or fall back to conventional way-partition for $B_2$.
$B_2$ can be made thread private if every thread makes
its own pack of the whole $B_1$ matrix.
The first choice achieves a full per-thread data isolation
at the expense of redundant packing overhead.
The second choice avoids the extra overhead in both time and space
but risks inter-thread conflicts caused by the shared $B_2$ matrix.
The two choices are evaluated and compared in Section~\ref{sec:evaluation}.

\subsection{Formal Description}\label{subsec:formal}
Given tiling factors $M_r$, $N_r$, $K_c$, $M_c$, $N_c$,
and architectural details of the memory hierachy,
The SCP method systematically determines the
the memory layout for packed matrices $A_2$ and $B_2$ used in GEMM.
We assume that the share matrix $B_2$ is packed in conventional way-partition style.
If desired, the $B_2$ can be privatized and handled in the same way as $A_2$.

@@

The memory layout of a packed matrix is described by a memory descriptor,
which is essentially a subspace of the whole address space $\mathcal{A}$.
Each packed matrix $M$ used by thread $t$,
is associated with a memory descriptor $\mathcal{D}_t^M$
specifying the memory space it occupies.
Generally, $\mathcal{D}_t^M$ consists of a sequence of $N_t^M$ disjoint memory segments,
$\mathcal{D}_t^M = \{ S_0, S_1, \cdots, S_{N_t^M}\}$.
%% in which $S_i \bigcap S_j = \phi$ if $i \ne j$.
A segment is represented by its start and end addresses $S_i = [s_i, e_i)$,
in which $s_i < e_i \in \mathcal{A}$.
With way-partition, the memory descriptor contains only one segment.
With set-partition used in previous section,
the memory descriptor contains a sequence of equally strided segments.

@@

SCP runs in three phases.
The first phase allocates memory buffers for all $A_2$ and $B_2$
instances used in GEMM. There are two buffers, $\mathcal{D}_p$ and $\mathcal{D}_s$,
for storing private and shared matrices respectively.
The second phase computes memory descriptors for $A_2$ instances
by partitioning $\mathcal{D}_p$ into $NT$ thread-private buffers
$\mathcal{D}_t^{A_2}$ , $t \in [0, NT)$.
Buffers of different threads are disjoint,
i.e. $\mathcal{D}_i^{A_2} \bigcap \mathcal{D}_j^{A_2} = \phi$ if $i \ne j$.
The third phase computes memory descriptors for $B_2$.
Because is $B_2$ is shared among all threads,
$\mathcal{D}_t^{B_2}$ does not represent the whole $B_2$ matrix,
but the part (size $\frac{es N_c K_c}{NT}$) which is packed by thread $t$.
The working flow of these three phases are listed
in Algorithm~\ref{alg:scp.phase1} -- \ref{alg:scp.phase3}.

In Algorithm~\ref{alg:scp.phase1},
the per-thread size for memory buffer $\mathcal{D}_p$ and
$\mathcal{D}_s$ are first computed
(line~\ref{line:size.p}--\ref{line:size.s}).
Because the memory space in $\mathcal{D}_p$ will be set-partitioned,
extra efforts are made to ensure a proper alignment for $\mathcal{D}_p$.
Line~\ref{line:align.init}--\ref{line:align.endfor} compute a value $align$,
and $size_p$ is enlarged to a multile of $align$ (line~\ref{line:align}).
$align$ is initialized with the L1 cache line size $l_1$.
Then on each level $l$ (Line~\ref{line:align.for})
with shared non-LRU caches (Line~\ref{line:align.type}),
$align$ is updated by computing the least-common-multiple of
original $align$ and $wc_l/nt_l$ (line~\ref{line:align.update}).
The insight in the $wc_l/nt_l$ value is that
each way of the level-$l$ cache is equally partitioned among $nt_l$ threads
in the set-partition of shared caches.
At last, memory are allocated (line~\ref{line:alloc.begin}--\ref{line:alloc.end})
and buffers $\mathcal{D}_p$ and $\mathcal{D}_s$
are created (line~\ref{line:d.begin}--\ref{line:d.end}).

@@

%% FIXME memory buffer? memory space? subspace?
Algorithm~\ref{alg:scp.phase2} shows the second phase of SCP.
The main part is a recursive procedure $subspace$
(line~\ref{line:subspace.begin}--\ref{line:subspace.end}).
The $subspace$ procedure traverses the memory hierachy
to compute for each node, including the processor cores and main memory,
a subspace of $\mathcal{D}_p$.
$subspace$ has three input parameters, $l$ and $idx$ to
identify the node on which it is running,
and a subspace $\mathcal{D}_l^{idx} \subset \mathcal{D}_p$ allocated to the node. 
The functionality of $subspace$ is to partition $\mathcal{D}_l^{idx}$
among all children of node $(l, idx)$.
If $subspace$ encounters a level 0 node, i.e. a processor core,
then $\mathcal{D}_l^{idx}$ is exactly the thread-private
memory buffer $\mathcal{D}_{idx}$ for thread $idx$ and $subspace$ returns early
(line~\ref{line:shortcut.begin}--\ref{line:shortcut.end}).
If this is not the case, the number of children is computed (line~\ref{line:nchildren})
and $subspace$ iterates over all the child nodes
(line~\ref{line:for.begin}--\ref{line:for.end}).
For each child node $cidx$ on layer $l-1$ (line~\ref{line:cidx}),
its memory space $\mathcal{D}_{l-1}^{cidx}$ is computed
and $subspace$ is called recursively on it (line~\ref{line:recursive}).
$\mathcal{D}_{l-1}^{cidx} \subset \mathcal{D}_l^{idx}$ is obtained
by intersecting $\mathcal{D}_l^{idx}$ with
a temporal memory space $\mathcal{D}_{temp}$ (line~\ref{line:childspace}).
The if-else branches (line~\ref{line:if}--\ref{line:endif})
show two distinct ways to compute $\mathcal{D}_{temp}$.
If the node is main memory or a LRU cache (line~\ref{line:if}),
$\mathcal{D}_l^{idx}$ is partitioned in the conventional
way-partition style.
Otherwise, the else branch (line~\ref{line:else}) performs
a set-partition on $\mathcal{D}_l^{idx}$.
Initially, the main memory gets the whole $\mathcal{D}_p$ (line~\ref{line:memory.d})
and $nt_{NL+1}$ is set to $NT$ (line~\ref{line:memory.nt})
because the main memory is shared by all threads.
Then $subspace$ procedure starts from the main memory (line~\ref{line:subspace.root})
and traverses the memory hierachy in a depth-first-search order.

The computation of $\mathcal{D}_t^{B_2}$ is quite simple.
Algorithm~\ref{alg:scp.phase3} divides $\mathcal{D}_s$
into $NT$ partitions, one for each thread.
First, the range of $\mathcal{D}_s$ is obtained
(line~\ref{line:lb}--\ref{line:ub}) and
the capacity of $\mathcal{D}_s$ is equally
divided among $NT$ threads (line~\ref{line:len}).
Then $\mathcal{D}_s$ is partitioned in a way-partition style
and each thread $t$ get buffer space for its part in $B_2$
(line~\ref{line:thread.for}--\ref{line:thread.forend}).

>>>>

This section demonstrates the SCP method to handle inter-thread cache conflicts.
The basic idea is that inter-thread cache conflicts can be eliminated
if packed matrices used by different threads are fetched to different cache sets.
In Section~\ref{subsec:example}, we demonstrate SCP by an example.
Then a formal description is given in Section~\ref{subsec:formal}.

@@

Suppose we are implementing a double precision GEMM (DGEMM) on
the 4-core processor, i.e. $es=8B$.
We now determine the tiling factors used in DGEMM following the procedure described
by constraints (\ref{eq:constraints.reg})--(\ref{eq:constraints.l3}).
Various configurations for ($M_r$, $N_r$) can satisfy (\ref{eq:constraints.reg}),
e.g. $(4,8)$, $(8,4)$, $(6,8)$, and $(8,6)$. 
It takes careful consideration to determine which one is the best,
and a tuning process may be needed.
After tuning we obtain the best values $M_r = 4$ and $N_r = 8$.
By (\ref{eq:constraints.l1}),
$K_c \le \lfloor c_{L1}/es/nt_{L1}/(N_r + 2 M_r) \rfloor = \lfloor 32KB/8B/1/16 \rfloor = 256$.
Here we choose $K_c=256$.
By (\ref{eq:constraints.l2}),
$M_c \le \lfloor (c_{L2}/es/nt_{L2} - 2 N_r K_c )/ K_c \rfloor =
\lfloor (2MB/8B/4 - 2*8*256)/256 \rfloor = 240$.
To leave space for other data structures, as well as the program code,
it is reasonable to shrink $M_c$ to a slightly smaller value,
e.g. $192$, $208$ and $224$ are all proper candidates for $M_c$.
Similar to $M_r$ and $N_r$, a tunning process may be needed to
find the best value for $M_c$.
Here we choose $M_c = 192$.
As there is no L3 cache, (\ref{eq:constraints.l3}) can be ignored
and $N_c$ can be given a large value.
Here we choose $N_c = 1024NT = 4096$.
With tiling factors determined, the sizes of packed matrices
can be calculated. Table.~\ref{tab:msizes} list the sizes of packed matrices.

@@

Figure.~\ref{fig:partition.conventional} and Figure.~\ref{fig:partition.segmented}
essentially represent two different partitioning style of the shared cache,
refered as way-partitioning and set-partitioning, respectively.
To guarantee the data layout in set-partitioning,
$A_2$ cannot be stored in a single continous buffer any more.
Instead, $A_2$ is distributed to 12 continous memory segments, each of size $32KB$.
The distance from one segment to the next is equal to $wc_2=128KB$.
%% There is no memory fragment because the segment size is a multiple of
%% size of $A_4$ (8KB).
%% FIXME put this to the end of this section
While the set-partitioning of shared caches avoids inter-thread cache conflicts,
it introduces some complexity to GEMM implementation
because the GEBP kernel and the matrix packing routines
now must work with segmented memory buffers instead of continous ones.

How about $B_2$, the other packed matrix use in GEBP?
Unlike $A_2$, which is thread private data,
$B_2$ is shared among all threads and
the packing of $B_2$ is done collabratively by all threads.
We have two choices, privatize $B_2$ and apply the set-partitioning method,
or fall back to conventional way-partitioning for $B_2$.
$B_2$ can be made thread private if every thread makes
its own pack of the whole $B_1$ matrix.
The first choice achieves a full per-thread data isolation
at the expense of redundant packing overhead.
The second choice avoids the extra overhead in both time and space
but risks inter-thread conflicts caused by the shared $B_2$ matrix.
The two choices are evaluated and compared in Section~\ref{sec:evaluation}.

\subsection{Formal Description}\label{subsec:formal}
Given tiling factors $M_r$, $N_r$, $K_c$, $M_c$, $N_c$,
and architectural details of the memory hierachy,
The SCP method systematically determines the
the memory layout for packed matrices $A_2$ and $B_2$ used in GEMM.
We assume that the share matrix $B_2$ is packed
in conventional way-partitioning style.
If desired, $B_2$ can be privatized and handled in the same way as $A_2$.

@@

The memory layout of a packed matrix is described by a memory descriptor,
which is essentially a subspace of the whole address space $\mathcal{A}$.
Each packed matrix $M$ used by thread $t$,
is associated with a memory descriptor $\mathcal{D}_t^M$
specifying the memory space it occupies.
Generally, $\mathcal{D}_t^M$ consists of a sequence of $N_t^M$ disjoint memory segments,
$\mathcal{D}_t^M = \{ S_0, S_1, \cdots, S_{N_t^M}\}$.
%% in which $S_i \bigcap S_j = \phi$ if $i \ne j$.
A segment is represented by its start and end addresses $S_i = [s_i, e_i)$,
in which $s_i < e_i \in \mathcal{A}$.
With way-partitioning, the memory descriptor contains only one segment.
With set-partitioning used in previous section,
the memory descriptor contains a sequence of equally strided segments.

@@

SCP runs in three phases.
The first phase allocates memory buffers for all $A_2$ and $B_2$
instances used in GEMM. There are two buffers, $\mathcal{D}_p$ and $\mathcal{D}_s$,
for storing private and shared matrices respectively.
The second phase computes memory descriptors for $A_2$ instances
by partitioning $\mathcal{D}_p$ into $NT$ thread-private buffers
$\mathcal{D}_t^{A_2}$ , $t \in [0, NT)$.
Buffers of different threads are disjoint,
i.e. $\mathcal{D}_i^{A_2} \bigcap \mathcal{D}_j^{A_2} = \phi$ if $i \ne j$.
The third phase computes memory descriptors for $B_2$.
Because is $B_2$ is shared among all threads,
$\mathcal{D}_t^{B_2}$ does not represent the whole $B_2$ matrix,
but the part (of size $\frac{es N_c K_c}{NT}$) which is packed by thread $t$.
The working flow of these three phases is listed
in Algorithm~\ref{alg:scp.phase1} -- \ref{alg:scp.phase3}.

In Algorithm~\ref{alg:scp.phase1},
the per-thread size for memory buffer $\mathcal{D}_p$ and
$\mathcal{D}_s$ are first computed
(line~\ref{line:size.p}--\ref{line:size.s}).
Because the memory space in $\mathcal{D}_p$ will be set-partitioned,
extra efforts are made to ensure a proper alignment for $\mathcal{D}_p$.
Line~\ref{line:align.init}--\ref{line:align.endfor} compute a value $align$,
and $size_p$ is enlarged to a multile of $align$ (line~\ref{line:align}).
$align$ is initialized with the L1 cache line size $l_1$.
Then on each level $l$ (line~\ref{line:align.for})
with shared non-LRU caches (line~\ref{line:align.type}),
$align$ is updated by computing the least-common-multiple of
original $align$ and $wc_l/nt_l$ (line~\ref{line:align.update}).
The insight in the $wc_l/nt_l$ value is that
each way of the level-$l$ cache is equally partitioned among $nt_l$ threads
in the set-partitioning of shared caches.
At last, memory is allocated (line~\ref{line:alloc.begin}--\ref{line:alloc.end})
and buffers $\mathcal{D}_p$ and $\mathcal{D}_s$
are created (line~\ref{line:d.begin}--\ref{line:d.end}).

@@

%% FIXME memory buffer? memory space? subspace?
Algorithm~\ref{alg:scp.phase2} shows the second phase of SCP.
The main part is a recursive procedure $subspace$
(line~\ref{line:subspace.begin}--\ref{line:subspace.end}).
The $subspace$ procedure traverses the memory hierachy
to compute for each node, including the processor cores and main memory,
a subspace of $\mathcal{D}_p$.
$subspace$ has three input parameters, $l$ and $idx$ to
identify the node on which it is running,
and a subspace $\mathcal{D}_l^{idx} \subset \mathcal{D}_p$ allocated to the node. 
The functionality of $subspace$ is to partition $\mathcal{D}_l^{idx}$
among all children of node $(l, idx)$.
If $subspace$ encounters a level 0 node, i.e. a processor core,
then $\mathcal{D}_l^{idx}$ is exactly the thread-private
memory buffer $\mathcal{D}_{idx}$ for thread $idx$ and $subspace$ returns early
(line~\ref{line:shortcut.begin}--\ref{line:shortcut.end}).
If this is not the case, the number of children is computed (line~\ref{line:nchildren})
and $subspace$ iterates over all the child nodes
(line~\ref{line:for.begin}--\ref{line:for.end}).
For each child node $cidx$ on layer $l-1$ (line~\ref{line:cidx}),
its memory space $\mathcal{D}_{l-1}^{cidx}$ is computed
and $subspace$ is called recursively on it (line~\ref{line:recursive}).
$\mathcal{D}_{l-1}^{cidx} \subset \mathcal{D}_l^{idx}$ is obtained
by intersecting $\mathcal{D}_l^{idx}$ with
a temporal memory space $\mathcal{D}_{temp}$ (line~\ref{line:childspace}).
The if-else branches (line~\ref{line:if}--\ref{line:endif})
show two distinct ways to compute $\mathcal{D}_{temp}$.
If the node is main memory or a LRU cache (line~\ref{line:if}),
$\mathcal{D}_l^{idx}$ is partitioned in the conventional
way-partitioning style.
Otherwise, the else branch (line~\ref{line:else}) performs
a set-partitioning on $\mathcal{D}_l^{idx}$.
Initially, the main memory gets the whole $\mathcal{D}_p$ (line~\ref{line:memory.d})
and $nt_{NL+1}$ is set to $NT$ (line~\ref{line:memory.nt})
because the main memory is shared by all threads.
Then $subspace$ procedure starts from the main memory (line~\ref{line:subspace.root})
and traverses the memory hierachy in a depth-first-search order.

The computation of $\mathcal{D}_t^{B_2}$ is quite simple.
Algorithm~\ref{alg:scp.phase3} divides $\mathcal{D}_s$
into $NT$ partitions, one for each thread.
First, the range of $\mathcal{D}_s$ is obtained
(line~\ref{line:lb}--\ref{line:ub}) and
the capacity of $\mathcal{D}_s$ is equally
divided among $NT$ threads (line~\ref{line:len}).
Then $\mathcal{D}_s$ is partitioned in a way-partitioning style
and each thread $t$ get buffer space for its part in $B_2$
(line~\ref{line:thread.for}--\ref{line:thread.forend}).
